Sleep, relative to wake, increases both veridical and false memory in the DRM paradigm: A registered report

### 2023-07-01 manuscript.tex
First working version producing most figures

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{caption}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Sleep, relative to wake, increases both veridical and false memory in the DRM paradigm: A registered report},
  pdfauthor={Matthew H.C. Mak, Alice O'Hagan, Aidan J Horner, M Gareth Gaskell},
  colorlinks=true,
  linkcolor={black},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Sleep, relative to wake, increases both veridical and false memory in the DRM paradigm: A registered report}
\author{Matthew H.C. Mak, Alice O'Hagan, Aidan J Horner, M Gareth Gaskell}
\date{Department of Psychology, University of York, UK}

@@

To address Research Question \#1, we first tested if any difference between the Sleep and Wake groups was significantly different from that between the AM- and PM-control groups (i.e., an interaction between Interval and Test Time). This is important because it allows us to rule out time-of-day effects. Then, we tested for the simple effect of Test Time (9AM vs.~9PM) within the Immediate and Delay groups. If there is (1) a significant Interval \(\times\) Test Time interaction and (2) a significant Test Time effect within the Delay groups (Sleep \textgreater{} Wake), we will be able conclude that sleep (but not time-of-day) increases false recall.

@@

The procedure of the study is summarised in Figure 1. The study was hosted on Gorilla (www.gorilla.sc; {[}\protect\hyperlink{ref-anwyl-irvine2020a}{9}{]}). A study phase took approximately 11 minutes. Here, participants first gave informed consent, completed a language/attention check, rated their level of sleepiness on the Stanford Sleepiness Scale (SSS; {[}\protect\hyperlink{ref-hoddes1973a}{22}{]}), and viewed 20 DRM wordlists.

@@

@@

The free recall data were pre-processed. The first step was to remove any duplicate responses. The second was to correct all obvious spelling and typing errors to the nearest English words, defined as Levenstein distance \(\leq\) 2 (e.g., *cigerette → cigarette). Responses with added or dropped inflectional suffixes (i.e., -s, -ed, -ing, adjectival -er) were corrected. Responses with derivational changes were considered as intrusions. For instance, one of the studied words is pollution; if a participant recalled pollutions, the plural suffix was dropped; however, if a participant recalled pollutant, this was considered as an intrusion.
\# Results of Pre-registered Analyses

@@

Payne et al. {[}\protect\hyperlink{ref-payne2009a}{6}{]} found that participants falsely recalled more critical lures post-sleep (vs.~post-wake). However, it is possible that participants simply had a greater tendency to put down more unseen words after sleep, not because sleep increases DRM false recall per se. Therefore, before addressing our key research questions, we checked if participants across groups were comparable in terms of their bias in producing unseen items. In Payne et al. {[}\protect\hyperlink{ref-payne2009a}{6}{]}, this bias was indexed via the number of intrusions (i.e., neither the studied nor the lure items), which was roughly equivalent between their Sleep and Wake groups {[}MSleep = 5.6 vs.~MWake = 6.2; p = .60{]} as well as between their AM and PM-control groups {[}MAM = 4.1 vs.~MPM = 4.1; p = .99{]}. To check if this is the case in our data, we used a 2 (Interval: Immediate vs.~Delay) \(\times\) 2 (Test Time: AM vs.~PM) Poisson regression. We chose Poisson regression, as opposed to ANOVA, because the intrusion data were count data, meaning that data distribution were right-skewed and hence unsuitable for ANOVA. Figure 2 summarises the number of intrusions in each group.

@@

The number of critical lures falsely recalled is summarised across groups in Figure 3.

A generalised linear mixed-effect model (GLMM) was fitted to the critical lure data on the item level (N of observations = 488 participants \(\times\) 20 critical lures). The dependent variable was binary: whether a critical lure was recalled or not (1 vs.~0). The fixed effects were the number of intrusions a participant produced, Interval (Immediate vs.~Delay), Test Time (AM vs.~PM), and an Interval by Test Time interaction. Interval and Test Time were coded using sum contrasts {[}\protect\hyperlink{ref-barr2019a}{26}{]}. The random-effect structure was determined by the ``buildmer'' package {[}\protect\hyperlink{ref-voeten2021a}{27}{]}, which automatically found the maximal model that was capable of converging using backward elimination (with the ``bobyqa'' optimiser). This means that model selection started from the maximal model, as justified by the experimental design {[}\protect\hyperlink{ref-barr2013a}{28}{]}. The model we reported and based our interpretation on was the most maximal model that was capable of converging (see the upper half of Table 4 for the final random-effect structure and model output).

@@

Figure 4 summarises the number of studied list words correctly recalled across groups.

@@

Bayes Factors were computed for (1) the Interval \(\times\) Test Time interaction in the false and veridical mixed-effect models above, and for the simple effects of Test Time within the (2) Immediate and (3) Delay groups. Following the procedures in Gilbert et al. {[}\protect\hyperlink{ref-gilbert2018a}{29}{]}, a Bayes Factor was computed using the Bayesian Information Criterion (BIC) approximation from two competing GLMMs. For instance, in computing the Bayes Factor for the Interval \(\times\) Test Time interaction, two models were needed: An alternative model containing the full fixed-effects structure (Intrusions + Interval + Test Time + Interval \(\times\) Test Time), and a null model lacking the interaction. To estimate the Bayes Factor, we used the formula e \(\Delta\)BIC10/2, where \(\Delta\)BIC10 is the BIC for the null model minus the BIC for the alternative model {[}\protect\hyperlink{ref-masson2011a}{30}--\protect\hyperlink{ref-wagenmakers2007a}{32}{]}. This produces a Bayes Factor10, which was interpreted with reference to Lee and Wagenmakers' {[}\protect\hyperlink{ref-wagenmakers2014a}{\textbf{wagenmakers2014a?}}{]} heuristics. The current BIC approximation method has the advantage of being a straightforward solution for mixed-effects models; however, its usage remains controversial as it is known to favour the simpler model (i.e., the null hypothesis; {[}\protect\hyperlink{ref-lindelov2018a}{31},\protect\hyperlink{ref-vandekerckhove2014a}{\textbf{vandekerckhove2014a?}},\protect\hyperlink{ref-weakliem1999a}{\textbf{weakliem1999a?}}{]}). Table 5 summarises the Bayes Factors derived from our mixed-effects models.

@@

In this exploratory analysis, we followed Diekelmann et al. {[}\protect\hyperlink{ref-diekelmann2010a}{34}{]} by focusing on the Delay groups and by performing a median split on our sample's adjusted veridical recall (Median = 5), classifying our participants as either high (\textgreater5; N = 117) or low ( \(\leq\) 5; N = 123) performers. We followed the analysis approach of Diekelmann et al. {[}\protect\hyperlink{ref-diekelmann2010a}{34}{]}, but since the number of critical lures a participant produced is count data, we used a 2 (Group: Sleep vs.~Wake) \(\times\) 2 (Adjusted veridical recall: High vs.~Low Performers) Poisson regression, instead of an ANOVA. This analysis showed no main effect of Group (\(\beta\) = -0.052, z = -0.494, p = .621) but a main effect of Adjusted veridical recall (\(\beta\) = -0.438, z = -3.766, p \textless{} .001), such that high performers (M = 3.17, SD = 2.71) tended to falsely recall more critical lures than low performers (M = 2.09, SD = 1.68) ({[}\protect\hyperlink{ref-thapar2001a}{\textbf{thapar2001a?}},\protect\hyperlink{ref-toglia1999}{\textbf{toglia1999?}}{]}, but see {[}\protect\hyperlink{ref-roediger2001a}{21},\protect\hyperlink{ref-stadler1999a}{35},\protect\hyperlink{ref-cann2011a}{\textbf{cann2011a?}}{]}). Furthermore, importantly, the interaction between Group and Adjusted veridical recall was not significant (\(\beta\) = 0.049, z = 0.30, p = .763). Despite this, we followed Diekelmann et al. {[}\protect\hyperlink{ref-diekelmann2010a}{34}{]} by comparing low performers in the Sleep and Wake groups. Contrary to their findings, emmeans showed that our low performers in both groups recalled essentially the same number of critical lures (MSleep = 2.09, SDSleep = 1.63 vs.~Mwake = 2.09, SDWake = 1.74; \(\beta\) = 0.133, z = 0.021, p = .982). In other words, this exploratory analysis found no evidence that the effect of sleep on false recall was modulated by veridical memory, casting doubt over the robustness and reliability of Diekelmann et al.'s {[}\protect\hyperlink{ref-diekelmann2010a}{34}{]} findings.

@@

Here, we asked whether recall probability of a lure (e.g., doctor) is predicted by the number of corresponding list items being recalled (e.g., nurse, hospital, sick), and if it does, whether it differs between the Sleep and Wake groups. These questions help shed light on the degree to which sleep increases lure recall via processes such as retrieval-induced generalisation or gist abstraction , as these processes may predict a different degree of interdependence between lure and veridical recall. If sleep (vs.~wake) promoted retrieval-induced generalisation, lure and veridical recall should become more strongly correlated with each other after sleep, because better veridical recall for a set of studied words may generalise to the corresponding critical lure (or vice versa). On the other hand, if sleep (vs.~wake) promoted gist abstraction, lure recall may become less related to memories for the corresponding list items, because theories of gist abstraction such as iOtA may predict that sleep would selectively boost the overlapping gist memory (i.e., the lure) but not necessarily the specific studied words.

@@

Across the Immediate and Delay groups, the number of intrusions and correct recall per list both had a main effect (ps \(\leq\) .001) such that they were positively correlated with lure recall. However, the effect of Test Time was not significant in either group (ps \textgreater{} .11). Finally, for the Correct recall per list \(\times\) Test Time interaction, it was significant in the Delay (z = -3.504, p \textless{} .001) but not the Immediate groups (z = -0.037, p = .971). To interpret the former, we used the R package, effects {[}\protect\hyperlink{ref-fox2022a}{\textbf{fox2022a?}}{]}, to visualise it (see Figure 6A) and plotted lure recall probability against veridical recall on a participant level (see Figure 6B).

@@

\textbf{Mixed evidence in the existing literature}. Our well-powered registered report had a total sample size of 488, with 120 in each of the Sleep and Wake groups. This far surpasses the sample sizes of prior `Sleep \(\times\) DRM' studies, where the median N was 27.6 per group. In our data, the effect of sleep on lure recall was only significant when controlling for differences in intrusion rates, and we checked the size of this sleep effect in an exploratory analysis, putting the estimate at Cohen's d = 0.274. This estimate is substantially lower than that from Newbury and Monaghan's {[}\protect\hyperlink{ref-newbury2019a}{1}{]} meta-analysis, which reported to be Cohen's d = +0.92 (95\% CI: 0.54, 1.30). The larger sample size and well-powered nature of our registered report contribute to a more precise estimation of the effect size, highlighting the need for cautious interpretation of prior evidence and the possibility that prior effect sizes may have been inflated due to, for example, publication bias and small sample sizes. Furthermore, given sleep had such a small effect on DRM false memory, it is perhaps not surprising that previous studies had produced mixed evidence. In light of these, we echo the view of a recent article {[}\protect\hyperlink{ref-cordi2021a}{\textbf{cordi2021a?}}{]} that future sleep research must prioritise robust methodologies (e.g., registered report and pre-registration) and larger sample sizes to enhance the reliability and generalisability of sleep-related memory effects {[}see also \protect\hyperlink{ref-nemeth2021a}{\textbf{nemeth2021a?}}{]}.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{caption}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Sleep, relative to wake, increases both veridical and false memory in the DRM paradigm: A registered report},
  pdfauthor={Matthew H.C. Mak, Alice O'Hagan, Aidan J. Horner, M. Gareth Gaskell},
  colorlinks=true,
  linkcolor={black},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Sleep, relative to wake, increases both veridical and false memory in the DRM paradigm: A registered report}
\author{Matthew H.C. Mak, Alice O'Hagan, Aidan J. Horner, M. Gareth Gaskell}
\date{Department of Psychology, University of York, UK}

@@

To address Research Question \#1, we first tested if any difference between the Sleep and Wake groups was significantly different from that between the AM- and PM-control groups (i.e., an interaction between Interval and Test Time). This is important because it allows us to rule out time-of-day effects. Then, we tested for the simple effect of Test Time (9AM vs.~9PM) within the Immediate and Delay groups. If there is (1) a significant Interval x Test Time interaction and (2) a significant Test Time effect within the Delay groups (Sleep \textgreater{} Wake), we will be able conclude that sleep (but not time-of-day) increases false recall.\footnote{Prior studies in the ‘Sleep $\times$ DRM’ literature (e.g., Fenn et al., 2009; Payne et al., 2009) conducted two separate statistical tests, one comparing Sleep vs. Wake, another comparing AM- vs. PM- controls. They then concluded that Sleep had an effect on DRM false memory beyond time-of-day effects when Test Time was significant in the Sleep vs. Wake comparison ($p$ < 0.05) but not in the AM vs. PM comparison ($p$ > 0.05). Unfortunately, however, this is not sufficient (Nieuwenhuis et al., 2011), as “the difference between ‘significant’ and ‘not significant’ is not itself statistically significant” (Gelmam \& Stern, 2006). Therefore, in order to rule out time-of-day effects, one needs to show that Sleep vs. Wake is significantly different from AM vs. PM-control. This can be captured by an Interval $\times$ Test Time interaction.}

@@

The procedure of the study is summarised in Figure \ref{fig:expprocedure}. The study was hosted on Gorilla (www.gorilla.sc; {[}\protect\hyperlink{ref-anwyl-irvine2020a}{9}{]}). A study phase took approximately 11 minutes. Here, participants first gave informed consent, completed a language/attention check, rated their level of sleepiness on the Stanford Sleepiness Scale (SSS; {[}\protect\hyperlink{ref-hoddes1973a}{22}{]}), and viewed 20 DRM wordlists.

@@

\begin{figure}

{\centering \includegraphics{Figures/expprocedure}

}

\caption{Experimental procedure.}\label{fig:expprocedure}
\end{figure}

@@

The free recall data were pre-processed. The first step was to remove any duplicate responses. The second was to correct all obvious spelling and typing errors to the nearest English words, defined as Levenstein distance \(\leq\) 2 (e.g., *cigerette → cigarette).\footnote{If a misspelt response has more than one nearest word, the response was considered as an intrusion.} Responses with added or dropped inflectional suffixes (i.e., -s, -ed, -ing, adjectival -er) were corrected. Responses with derivational changes were considered as intrusions. For instance, one of the studied words is pollution; if a participant recalled pollutions, the plural suffix was dropped; however, if a participant recalled pollutant, this was considered as an intrusion.
\# Results of Pre-registered Analyses

@@

Payne et al. {[}\protect\hyperlink{ref-payne2009a}{6}{]} found that participants falsely recalled more critical lures post-sleep (vs.~post-wake). However, it is possible that participants simply had a greater tendency to put down more unseen words after sleep, not because sleep increases DRM false recall per se. Therefore, before addressing our key research questions, we checked if participants across groups were comparable in terms of their bias in producing unseen items. In Payne et al. {[}\protect\hyperlink{ref-payne2009a}{6}{]}, this bias was indexed via the number of intrusions (i.e., neither the studied nor the lure items), which was roughly equivalent between their Sleep and Wake groups {[}MSleep = 5.6 vs.~MWake = 6.2; p = .60{]} as well as between their AM and PM-control groups {[}MAM = 4.1 vs.~MPM = 4.1; p = .99{]}. To check if this is the case in our data, we used a 2 (Interval: Immediate vs.~Delay) \(\times\) 2 (Test Time: AM vs.~PM) Poisson regression. We chose Poisson regression, as opposed to ANOVA, because the intrusion data were count data, meaning that data distribution were right-skewed and hence unsuitable for ANOVA. Figure \ref{fig:intrusionfig} summarises the number of intrusions in each group.

\begin{figure}

{\centering \includegraphics{Figures/intrusiongraph}

}

\caption{Number of intrusions produced, summarised across the four groups. Each dot represents an individual participant, and the error bars represent 95\% confidence intervals.}\label{fig:intrusionfig}
\end{figure}

@@

The number of critical lures falsely recalled is summarised across groups in Figure \ref{fig:lurefig}.

\begin{figure}

{\centering \includegraphics{Figures/luregraph}

}

\caption{Number of critical lures falsely recalled, summarised across the four groups. Each dot represents an individual participant, and the error bars represent 95\% confidence intervals.}\label{fig:lurefig}
\end{figure}

A generalised linear mixed-effect model (GLMM) was fitted to the critical lure data on the item level (N of observations = 488 participants \(\times\) 20 critical lures).\footnote{The use of GLMM is a clear departure from prior ‘Sleep $\times$ DRM’ studies, the majority of which addressed the same research question using an independent t-test or ANOVA (e.g., Diekelmann et al., 2010; Monaghan et al., 2017; Payne et al., 2009). We explained in Appendix C why these statistical tests are usually not appropriate in the context of DRM recall and why GLMM are more advantageous.} The dependent variable was binary: whether a critical lure was recalled or not (1 vs.~0). The fixed effects were the number of intrusions a participant produced, Interval (Immediate vs.~Delay), Test Time (AM vs.~PM), and an Interval by Test Time interaction. Interval and Test Time were coded using sum contrasts {[}\protect\hyperlink{ref-barr2019a}{26}{]}. The random-effect structure was determined by the ``buildmer'' package {[}\protect\hyperlink{ref-voeten2021a}{27}{]}, which automatically found the maximal model that was capable of converging using backward elimination (with the ``bobyqa'' optimiser). This means that model selection started from the maximal model, as justified by the experimental design {[}\protect\hyperlink{ref-barr2013a}{28}{]}. The model we reported and based our interpretation on was the most maximal model that was capable of converging (see the upper half of Table 4 for the final random-effect structure and model output).

@@

Figure \ref{fig:studiedfig} summarises the number of studied list words correctly recalled across groups.

\begin{figure}

{\centering \includegraphics{Figures/studiedgraph}

}

\caption{Number of studied list words correctly recalled, summarised across the four groups. Each dot represents an individual participant, and the error bars represent 95\% confidence intervals.}\label{fig:studiedfig}
\end{figure}

@@

Bayes Factors were computed for (1) the Interval \(\times\) Test Time interaction in the false and veridical mixed-effect models above, and for the simple effects of Test Time within the (2) Immediate and (3) Delay groups. Following the procedures in Gilbert et al. {[}\protect\hyperlink{ref-gilbert2018a}{29}{]}, a Bayes Factor was computed using the Bayesian Information Criterion (BIC) approximation from two competing GLMMs. For instance, in computing the Bayes Factor for the Interval \(\times\) Test Time interaction, two models were needed: An alternative model containing the full fixed-effects structure (Intrusions + Interval + Test Time + Interval \(\times\) Test Time), and a null model lacking the interaction.\footnote{To obtain the Bayes Factor for the simple effects of Test Time, the alternative model will contain Test Time as the sole fixed effect while the null model will contain no fixed effects.} To estimate the Bayes Factor, we used the formula e \(\Delta\)BIC10/2, where \(\Delta\)BIC10 is the BIC for the null model minus the BIC for the alternative model {[}\protect\hyperlink{ref-masson2011a}{30}--\protect\hyperlink{ref-wagenmakers2007a}{32}{]}. This produces a Bayes Factor10, which was interpreted with reference to Lee and Wagenmakers' {[}\protect\hyperlink{ref-wagenmakers2014a}{\textbf{wagenmakers2014a?}}{]} heuristics. The current BIC approximation method has the advantage of being a straightforward solution for mixed-effects models; however, its usage remains controversial as it is known to favour the simpler model (i.e., the null hypothesis; {[}\protect\hyperlink{ref-lindelov2018a}{31},\protect\hyperlink{ref-vandekerckhove2014a}{\textbf{vandekerckhove2014a?}},\protect\hyperlink{ref-weakliem1999a}{\textbf{weakliem1999a?}}{]}). Table 5 summarises the Bayes Factors derived from our mixed-effects models.

@@

In this exploratory analysis, we followed Diekelmann et al. {[}\protect\hyperlink{ref-diekelmann2010a}{34}{]} by focusing on the Delay groups\footnote{Note that Diekelmann et al. (2010) did not have Immediate/Control groups.} and by performing a median split on our sample's adjusted veridical recall (Median = 5), classifying our participants as either high (\textgreater5; N = 117) or low ( \(\leq\) 5; N = 123) performers. We followed the analysis approach of Diekelmann et al. {[}\protect\hyperlink{ref-diekelmann2010a}{34}{]}, but since the number of critical lures a participant produced is count data, we used a 2 (Group: Sleep vs.~Wake) \(\times\) 2 (Adjusted veridical recall: High vs.~Low Performers) Poisson regression, instead of an ANOVA. This analysis showed no main effect of Group (\(\beta\) = -0.052, z = -0.494, p = .621) but a main effect of Adjusted veridical recall (\(\beta\) = -0.438, z = -3.766, p \textless{} .001), such that high performers (M = 3.17, SD = 2.71) tended to falsely recall more critical lures than low performers (M = 2.09, SD = 1.68) ({[}\protect\hyperlink{ref-thapar2001a}{\textbf{thapar2001a?}},\protect\hyperlink{ref-toglia1999}{\textbf{toglia1999?}}{]}, but see {[}\protect\hyperlink{ref-roediger2001a}{21},\protect\hyperlink{ref-stadler1999a}{35},\protect\hyperlink{ref-cann2011a}{\textbf{cann2011a?}}{]}). Furthermore, importantly, the interaction between Group and Adjusted veridical recall was not significant (\(\beta\) = 0.049, z = 0.30, p = .763). Despite this, we followed Diekelmann et al. {[}\protect\hyperlink{ref-diekelmann2010a}{34}{]} by comparing low performers in the Sleep and Wake groups. Contrary to their findings, emmeans showed that our low performers in both groups recalled essentially the same number of critical lures (MSleep = 2.09, SDSleep = 1.63 vs.~Mwake = 2.09, SDWake = 1.74; \(\beta\) = 0.133, z = 0.021, p = .982). In other words, this exploratory analysis found no evidence that the effect of sleep on false recall was modulated by veridical memory, casting doubt over the robustness and reliability of Diekelmann et al.'s {[}\protect\hyperlink{ref-diekelmann2010a}{34}{]} findings.

@@

Here, we asked whether recall probability of a lure (e.g., doctor) is predicted by the number of corresponding list items being recalled (e.g., nurse, hospital, sick), and if it does, whether it differs between the Sleep and Wake groups. These questions help shed light on the degree to which sleep increases lure recall via processes such as retrieval-induced generalisation\footnote{Retrieval-induced generalisation: Retrieval of one word cueing retrieval of a related word (e.g., Berens \& Bird, 2017)} or gist abstraction\footnote{Gist abstraction: Extraction of the central or essential meaning of learned information}, as these processes may predict a different degree of interdependence between lure and veridical recall. If sleep (vs.~wake) promoted retrieval-induced generalisation, lure and veridical recall should become more strongly correlated with each other after sleep, because better veridical recall for a set of studied words may generalise to the corresponding critical lure (or vice versa). On the other hand, if sleep (vs.~wake) promoted gist abstraction, lure recall may become less related to memories for the corresponding list items, because theories of gist abstraction such as iOtA may predict that sleep would selectively boost the overlapping gist memory (i.e., the lure) but not necessarily the specific studied words.

@@

Across the Immediate and Delay groups, the number of intrusions and correct recall per list both had a main effect (ps \(\leq\) .001) such that they were positively correlated with lure recall. However, the effect of Test Time was not significant in either group (ps \textgreater{} .11). Finally, for the Correct recall per list \(\times\) Test Time interaction, it was significant in the Delay (z = -3.504, p \textless{} .001) but not the Immediate groups (z = -0.037, p = .971). To interpret the former, we used the R package, effects {[}\protect\hyperlink{ref-fox2022a}{\textbf{fox2022a?}}{]}, to visualise it (see Figure \ref{fig:delayfig}A) and plotted lure recall probability against veridical recall on a participant level (see Figure \ref{fig:delayfig}B).

\begin{figure}

{\centering \includegraphics{Figures/delaygraph}

}

\caption{(A) Prediction from generalised mixed-effect model on the combined effects of correct recall/list and group (sleep vs. wake) on false recall (B) Correlation between mean lure and studied word recall in the Sleep and Wake groups. Each dot represents an individual participant. Dotted lines/shaded areas represent 95\% confidence intervals.}\label{fig:delayfig}
\end{figure}

@@

\textbf{Mixed evidence in the existing literature}. Our well-powered registered report had a total sample size of 488, with 120 in each of the Sleep and Wake groups. This far surpasses the sample sizes of prior `Sleep \(\times\) DRM' studies, where the median N was 27.6 per group. In our data, the effect of sleep on lure recall was only significant when controlling for differences in intrusion rates, and we checked the size of this sleep effect in an exploratory analysis, putting the estimate at Cohen's d = 0.274.\footnote{We used the eff\_size function in the emmeans package to estimate the effect size of Test Time (sleep vs. wake) in our confirmatory GLM model, with sigma = sigma(lure\_model) and edf = infinite.} This estimate is substantially lower than that from Newbury and Monaghan's {[}\protect\hyperlink{ref-newbury2019a}{1}{]} meta-analysis, which reported to be Cohen's d = +0.92 (95\% CI: 0.54, 1.30). The larger sample size and well-powered nature of our registered report contribute to a more precise estimation of the effect size, highlighting the need for cautious interpretation of prior evidence and the possibility that prior effect sizes may have been inflated due to, for example, publication bias and small sample sizes. Furthermore, given sleep had such a small effect on DRM false memory, it is perhaps not surprising that previous studies had produced mixed evidence. In light of these, we echo the view of a recent article {[}\protect\hyperlink{ref-cordi2021a}{\textbf{cordi2021a?}}{]} that future sleep research must prioritise robust methodologies (e.g., registered report and pre-registration) and larger sample sizes to enhance the reliability and generalisability of sleep-related memory effects {[}see also \protect\hyperlink{ref-nemeth2021a}{\textbf{nemeth2021a?}}{]}.
